Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-04-15 article.tex
Intro progress

\richel{Check for correctness}
Our immune system fights invaders on a daily basis.
These invaders can be fungi, bacteria or viruses.
The innate immune response is its first general 
and immediate strategy, where the acquired immune response
needs time to develop its specialized and more effective
combat forces.

@@

\richel{Check for correctness}
All nucleated cells in humans present randomly sampled polypeptide
fragments to the surroundings of the cell using MHC class I molecules.
If a cell gets infected by a virus, also the virus' polypeptides
will be presented at the cell surface. The viral antigen is detected 
by cytotoxic T lymphocytes, which will kill the infected cell.

@@

\richel{Check for correctness}
Bacteria and fungi do not infect a cell with their foreign DNA,
but are detected by the proteins on their cell walls. MHC class II molecules 
bind to specific polypeptide fragments. This binding allows T and B cells,
which all express MHC-II, to detect a pathogen and trigger the response to
kill the pathogen. Note that the foreign proteins of a viral envelope 
can be detected in this way as well, which is commonly used in vaccination.

@@

\richel{Check for correctness}
Any human's immune system detects only a selection of all possible
polypeptide fragments. Each MHC molecule can only express (MHC-I)
or detect (MHC-II) a couple of polypeptides. An increased detection
is obtained by expressing a wide variety of
MHCs, which is assured by the human leukocyte antigen gene complex (HLA).
However, a different HLA will result in a different variety of MHCs,
that will display different polypeptide fragments. Additionally, there
may have been selection the HLA to display and or detect different
polypeptide fragments.

@@

For MHC-I, it was found that predicted epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all human leukocyte antigen (HLA)-A and 
most HLA-B super types (\cite{bianchi2017}). 
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs have a reduced variability 
due to the functional requirement of being able to span a lipid 
bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.

@@

\paragraph{Evolution}

If the human immune system is more prone to detect polypeptides
derived from a pathogen's TMHs, one may expect to find an evolutionary
signal in fast-evolving pathogen. 
We can assume that only after the moment that COVID-19 spilled over 
to humans, it's the human immune system that is an influential component 
of its evolutionary selection.
Using the extensive amount of time-dated COVID-19 sequences,
we can quantify to which extent selection in COVID-19 occurs
in the transmembrane part of its proteome.

>>>>

\richel{Check against stupidity}
Our immune system fights invaders on a daily basis.
These invaders can be fungi, bacteria or viruses.
The innate immune response is its first general 
and immediate strategy, where the acquired immune response
needs time to develop its specialized and more effective
combat forces.

@@

\richel{Check against stupidity}
All nucleated cells in humans present randomly sampled polypeptide
fragments to the surroundings of the cell using MHC class I molecules.
If a cell gets infected by a virus, also the virus' polypeptides
will be presented at the cell surface. The viral antigen is detected 
by cytotoxic T lymphocytes, which will kill the infected cell.

@@

\richel{Check against stupidity}
Bacteria and fungi do not infect a cell with their foreign DNA,
but are detected by the proteins on their cell walls. MHC class II molecules 
bind to specific polypeptide fragments. This binding allows T and B cells,
which all express MHC-II, to detect a pathogen and trigger the response to
kill the pathogen. Note that the foreign proteins of a viral envelope 
can be detected in this way as well, which is commonly used in vaccination.

@@

\richel{Check against stupidity}
Any human's immune system detects only a selection of all possible
polypeptide fragments. Each MHC molecule can only express (MHC-I)
or detect (MHC-II) a couple of polypeptides. An increased detection
is obtained by expressing a wide variety of
MHCs, which is assured by the human leukocyte antigen gene complex (HLA).
However, a different HLA will result in a different variety of MHCs,
that will display different polypeptide fragments. Additionally, there
may have been selection the HLA to display and or detect different
polypeptide fragments.

@@

For MHC-I, it was found that predicted epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all HLA-A and 
most HLA-B super types (\cite{bianchi2017}). 
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs have a reduced variability 
due to the functional requirement of being able to span a lipid 
bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.

@@

\paragraph{Evolutionary arms race}

\richel{Check against stupidity}
The human immune system and human pathogen are in an evolutionary
arms race, in which our immune systems is selected for the detection
of pathogens, whereas pathogens are selected to avoid being detected.
If the human immune system is more prone to display (and possibly 
detect) polypeptides derived from TMHs, one may expect to find 
an evolutionary signal in pathogens as well.

\paragraph{Evolutionary signal of unknown strength}

\richel{Check against stupidity}
We do not know how viruses are selected \richel{so I should read the
literature on that}. We predict that there is a strong 
evolutionary conservation in regions that are essential for viral proteins'
functions. Because these regions are rare in a viral proteome,
and differ widely between pathogens, we expect that the human immune system
is not tailored to detect these sites. TMHs, however, have less 
functional constraints yet are more general. Because TMHs are more general,
we expect it for the immune system to pay off to be more attentive towards
TMHs. Because the strength of the evolutionary signal,
are strong data set needs to be used.

We expect that

We can assume that only after the moment that COVID-19 spilled over 
to humans, it's the human immune system that is an influential component 
of its evolutionary selection.
Using the extensive amount of time-dated COVID-19 sequences,
we can quantify to which extent selection in COVID-19 occurs
in the transmembrane part of its proteome.
